---
title: "Lab 5"
author: "Gage Clawson"
date: "2/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. attach packages

```{r}
library(tidyverse)
library(tseries)
library(forecast)

```

b. Get data

US Residential energy consumption values (jan 1973 - oct 2017), Trillion BTU (US Energy Information Administration)

```{r}
energy <- read_csv("energy.csv")
```

c. Create 'ts' timeseries data using information in the res_total column from energy

```{r}
res_ts <- ts(energy$res_total, frequency = 12, start = c(1973,1))
res_ts

plot(res_ts)
```

d. Decompose ts

```{r}
res_dc <- decompose(res_ts)
plot(res_dc)
```

e. Other cool ways to present ts data

```{r}
monthplot(res_ts)
```

```{r}
ggseasonplot(res_ts, polar= TRUE) +
  theme_dark()

```

f. Simple moving averages

forecast::ma()

```{r}
sma_res <- ma(res_ts, order = 5)
plot(res_ts)
lines(sma_res, col = "red")
```

g. ACF (autocorrelation)

```{r}
res_acf <- acf(res_ts)

ggtsdisplay(res_ts)
```

i. Holt-Winters exponential smoothing
```{r}
res_hw <- HoltWinters(res_ts)
res_hw

plot(res_hw)
```

Now some future forecasting with holt winters

```{r}
res_forecast <- forecast(res_hw, h = 60)

plot(res_forecast)
```

Check distribution of residuals 

```{r}
hist(res_forecast$residuals)
res_forecast
```

Autoregressive Intergrated Moving Average (ARIMA) using auto.arima for p,d,q

```{r}
res_pdq <- auto.arima(res_ts)

res_pdq
```

pdq: (1,0,2), (0,1,1)

fit the ARIMA model using the pdq inputs

```{r}

res_arima <- arima(res_ts, order = c(1,0,2), seasonal = list(order = c(0,1,1)))

res_arima

# Check out the residuals 
par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)


```


Now forecast with ARIMA outputs

```{r}
forecast_res <- forecast(res_arima, h = 72)
plot(forecast_res)
```
Now let's try viewing this from ggplot
```{r}
res_df <- data.frame(forecast_res)
month_seq = seq(1,72)

res_df2 <- data.frame(month_seq, res_df)

ggplot(res_df2, aes(x = month_seq, y = Point.Forecast)) +
  geom_line()
```







